Semi-Permeable Vesicles: Mathematical Formulation, Numerics, and Physics

### 2020-12-11 vesicle_code/docs/permeableVes/main.tex
yny

\dropcap{W}ater balance is an essential process in the adaptation of a living cell to its environment \cite{AlbertsMolecularBiology}.  For example, erythrocytes participate in efficient body water homeostasis by regulating its exchange of water  \cite{SugieIntaglietta2018_AmJPhysiolHCP}.  Another example is the simultaneous regulation of cell volume and the cell membrane tension during cell migration through trafficking of aquaporin proteins \cite{SaadounPapadopoulosWatanabeEtAl2005_JCS}.

The cell membrane is permeable to water and water-soluble molecules \cite{Dick1964_JTB,FettiplaceHaydon1980_PhysRev,DeamerBramhall1986_ChemPhysLipids,Grafmueller2019_ABLS}. In the presence of an osmotic gradient across the membrane (such as a contrast in the concentration of an impermeable solute), water flows in or out of the membrane to adjust the interior cellular volume. In the red blood cell membrane, aquaporin-1 (AQP1) channels give rise to a much higher water permeability of $\sim 1.8\times 10^{-2}$ cm/s in erythrocytes \cite{YangMaVerkman2001_JBC} while the intrinsic water permeability of a lipid bilayer membrane is in the range of $10^{-4}-10^{-3}$ cm/s \cite{ThompsonHuang1966_ANYAS,FettiplaceHaydon1980_PhysRev,Grafmueller2019_ABLS,Dimova2020_GVB}.

@@

\section*{Governing equations} \label{sc:formulate}
We consider a two-dimensional semi-permeable vesicle suspended in an
unbounded viscous fluid driven by the imposed flow $\uu_\infty(\xx)$.
Since fluid is transported across the semi-permeable vesicle membrane,
we only consider fluids with uniform viscosity. Letting $\gamma$ be the
vesicle interface, the fluid velocity $\uu$ and pressure $p$ satisfy
%
\begin{alignat}{3}
  \label{eqn:governing}
  -\nabla p + \mu \triangle \uu &= 0, \quad
  \nabla \cdot \uu = 0, \qquad &&\xx \in \RR^2 \setminus \gamma, \\
%
  \uu(\xx) &\rightarrow \uu_\infty(\xx),
    &&\norm{\xx} \rightarrow \infty,
\end{alignat}
where $\mu$ is the fluid viscosity. Letting $\nn$ be the outward normal
to $\gamma$, $\jump{\cdot}$ be the jump across the interface, and $T$ be
the hydrodynamic stress tensor, the interfacial boundary conditions are
\begin{align}
  \label{eqn:jump}
  \jump{T\cdot\nn} = \ff_\mathrm{mem}, \quad \jump{\uu} = 0,
\end{align}
where $\ff_\mathrm{mem}$ is the membrane force that results from the
Helfrich energy with bending modulus $k_b$, and a tension term that acts
as a Lagrange multiplier to satisfy the local inextensibility condition.
Following~\cite{vee-gue-zor-bir2009}, we modify the bending tension and
forces as
$\ff_\mathrm{mem} = \ff_\mathrm{ben} +
\ff_\mathrm{ten}$, where
\begin{align}
  \label{eqn:membraneForces}
  \ff_\mathrm{ben} = -k_b \xx_{ssss}, \qquad
  \ff_\mathrm{ten} = (\sigma \xx_s)_s.
\end{align}
The physical tension is $\Lambda = \sigma + 1.5 \kappa^2$, where
$\kappa$ is the membrane curvature.

>>>>

\dropcap{W}ater balance is an essential process in the adaptation of a living cell to its environment \cite{AlbertsMolecularBiology}.  Erythrocytes are known to  participate in efficient body water homeostasis by regulating its exchange of water  \cite{YangMaVerkman2001_JBC,SugieIntaglietta2018_AmJPhysiolHCP}.  Cells achieve motility by simultaneously regulating their cell volume and cell membrane tension through trafficking of aquaporin proteins \cite{SaadounPapadopoulosWatanabeEtAl2005_JCS,Verkman2008_JMM,BerthaudEtAl2016_SM} and flattening of caveolae \cite{Keren2011_EurBJ}. 
Water transport in and out of cell is also found to be important to 
active cell shape fluctuations leading to bleb formation \cite{TaloniKardashSalmanEtAl2015_PRL}.

The cell membrane is permeable to water and water-soluble molecules \cite{Dick1964_JTB,FettiplaceHaydon1980_PhysRev,DeamerBramhall1986_ChemPhysLipids,Grafmueller2019_ABLS}. In the presence of an osmotic gradient across the membrane (such as a contrast in the concentration of an impermeable solute), water flows in or out of the membrane to neutralize the osmotic gradient. In the red blood cell membrane, aquaporin-1 (AQP1) channels give rise to a much higher water permeability of $\sim 1.8\times 10^{-2}$ cm/s in erythrocytes \cite{YangMaVerkman2001_JBC} while the intrinsic water permeability of a lipid bilayer membrane is in the range of $10^{-4}-10^{-3}$ cm/s \cite{ThompsonHuang1966_ANYAS,FettiplaceHaydon1980_PhysRev,Grafmueller2019_ABLS,Dimova2020_GVB}.

@@

\section*{Governing equations \label{sc:formulate}}
We consider a two-dimensional semi-permeable vesicle suspended in an
unbounded viscous fluid driven by the imposed flow $\uu_\infty(\xx)$.
Since fluid is transported across the semi-permeable vesicle membrane,
we only consider fluids with uniform viscosity. Letting $\gamma$ be the
vesicle interface, the fluid velocity $\uu$ and pressure $p$ satisfy
%
\begin{alignat}{3}
  \label{eqn:governing}
  -\nabla p + \mu \triangle \uu &= 0, \quad
  \nabla \cdot \uu = 0, \qquad &&\xx \in \RR^2 \setminus \gamma, \\
%
  \uu(\xx) &\rightarrow \uu_\infty(\xx),
    &&\norm{\xx} \rightarrow \infty,
\end{alignat}
where $\mu$ is the fluid viscosity. Letting $\nn$ be the outward normal
to $\gamma$, $\jump{\cdot}$ be the jump across the interface, and $T$ be
the hydrodynamic stress tensor, the interfacial boundary conditions are
\begin{align}
  \label{eqn:jump}
  \jump{T\cdot\nn} = \ff_\mathrm{mem}, \quad \jump{\uu} = 0,
\end{align}
where $\ff_\mathrm{mem}$ is the membrane force that results from the
Helfrich energy with bending modulus $k_b$, and a tension term that acts
as a Lagrange multiplier to satisfy the local inextensibility condition.
Following~\cite{vee-gue-zor-bir2009}, we modify the bending tension and
forces as
$\ff_\mathrm{mem} = \ff_\mathrm{ben} +
\ff_\mathrm{ten}$, where
\begin{align}
  \label{eqn:membraneForces}
  \ff_\mathrm{ben} = -k_b \xx_{ssss}, \qquad
  \ff_\mathrm{ten} = (\sigma \xx_s)_s.
\end{align}
The physical tension is $\Lambda = \sigma + 1.5 \kappa^2$, where
$\kappa$ is the membrane curvature.
